Estimation of Manning's Friction Coefficient using Polynomial Chaos and Bayesian inference

### 2013-12-23 paper/coefs.tex
revisiting results section
modified intro and added one section on manning

We seek to infer the Manning coefficient from water surface elevation
data measured at different gauges as shown in Figure~\ref{fig:setup}.
Bayesian inference can be applied directly to infer the uncertain parameters
using Equation~\eqref{eq:post}. In our case, the observation data $\vec d$ 
are the water surface elevation at the gauges;
their model predicted counterparts $\vec G$ is predicted by \geoclaw.
The uncertain parameters $\vec \theta$ will be denoted by $\{N_i\}_{i=1}^3$ for the
three Manning Friction coefficients treated as the unknown parameters. 
In this case  Equation~\eqref{eq:post} can be written as:
\begin{equation} 
p(N_1,N_2,N_3,\sigma^2 | \vec d) 
\propto \frac{1}{\sqrt{2 \pi \sigma^2}} 
 \prod_{i=1}^m  
\exp \left\lbrace \frac{-(d_i - G_i)^2}{2 \sigma^2} \right\rbrace
\ p(\sigma^2)p(N_1)p(N_2) p(N_3)
\label{eq:post_coef}
\end{equation}

To complete the definition of the posterior, we need to choose a proper prior that should be based 
on some \emph{a priori} knowledge about the parameters. In our case, a uniform
prior for the model parameters is assumed as indicated:
\begin{equation} 
p(N_i) = \begin{cases}
		\displaystyle \frac{1}{b_i-a_i} &\text{for~} a_i <  N_i \leq b_i ,  \\
		0 &\text{otherwise}  , 
\end{cases}
\end{equation}
where $ [a_i,b_i]$ denote the parameter ranges defined as.
Regarding the variance, the only information we know 
is that $\sigma^2$ is always positive.
We thus assume a Jeffreys prior \citep{sivia}, expressed as:
\begin{equation} 
p(\sigma^2) =  \begin{cases}
		\displaystyle \frac{1}{\sigma^2} &\text{for~} \sigma^2 > 0,  \\
		0 &\text{otherwise}. 
		\end{cases}
\label{eq:var_pr}
\end{equation}

>>>>

We seek to infer the Manning's n coefficients from water surface elevation
data measured at different gauges as shown in Figure~\ref{fig:setup}.
Bayesian inference can be applied directly to infer the uncertain parameters
using Equation~\eqref{eq:post}. In our case, the observation data $\vec d$ 
are the water surface elevation at the gauges;
their model predicted counterparts $\vec G$ is predicted by \geoclaw.
The uncertain parameters $\vec \theta$ will be denoted by $\{N_i\}_{i=1}^3$ for the
three Manning's coefficients treated as the unknown parameters. 
In this case  Equation~\eqref{eq:post} can be written as:

\begin{equation} 
p(N_1,N_2,N_3,\sigma^2 | \vec d) 
\propto \frac{1}{\sqrt{2 \pi \sigma^2}} 
 \prod_{i=1}^m  
\exp \left\lbrace \frac{-(d_i - G_i)^2}{2 \sigma^2} \right\rbrace
\ p(\sigma^2)p(N_1)p(N_2) p(N_3)
\label{eq:post_coef}
\end{equation}

To complete the definition of the posterior, we need to choose a proper prior that should be based 
on some \emph{a priori} knowledge about the parameters. In our case, a uniform
prior for the model parameters is assumed as indicated:

\begin{equation} 
p(N_i) = \begin{cases}
		\displaystyle \frac{1}{b_i-a_i} &\text{for~} a_i <  N_i \leq b_i ,  \\
		0 &\text{otherwise}  , 
\end{cases}
\end{equation}
where $ [a_i,b_i]$ denote the parameter ranges defined as.
Regarding the variance, the only information we know 
is that $\sigma^2$ is always positive.
We thus assume a Jeffreys prior \citep{sivia}, expressed as:

\begin{equation} 
p(\sigma^2) =  \begin{cases}
		\displaystyle \frac{1}{\sigma^2} &\text{for~} \sigma^2 > 0,  \\
		0 &\text{otherwise}. 
		\end{cases}
\label{eq:var_pr}
\end{equation}

### 2013-12-23 paper/formulation.tex
modified intro and added one section on manning

\section{Formulation}
\label{sec:formu}
In this section, we introduce the methods employed in this work.
First, the Bayesian approach to inverse problems is described in Section~\ref{sec:inference};
its application to our problem of inferring the Manning Friction Coefficients in Section~\ref{sec:manning}.
Finally, in Section~\ref{sec:uqpce}, we introduce
a stochastic spectral method for forward uncertainty propagation and accelerating the Bayesian inference.

>>>>

\section{Formulation}
\label{sec:formu}
In this section, we introduce the methods employed in this work.
First, the Bayesian approach to inverse problems is described in Section~\ref{sec:inference};
its application to our problem of inferring the Manning's n friction coefficients in Section~\ref{sec:manning}.
Finally, in Section~\ref{sec:uqpce}, we introduce
a stochastic spectral method for forward uncertainty propagation and accelerating the Bayesian inference.

### 2013-12-23 paper/intro.tex
last changes from my side... i think
modified intro and added one section on manning

Tsunamis in the past decade have been responsible for some of the most deadly and costly natural disasters ever recorded.  Coastal communities have faced this hazard by assessing the risk of these hazards, attempting to make informed decisions about the likelihood that such an event would occur with cost to life and property.  To do this, computational models of tsunami events are often employed to explore multiple different scenarios.  Unfortunately, since tsunami events are relatively rare, there is an abundance of uncertainty in the input data for these computational models ranging from effects of the domain such as bathymetry and friction parameterizations, to the earthquake source.  In this article we present an approach to quantifying the uncertainty in the friction field and a method to find the best parameterized values of the friction field using the \tohoku tsunami as a case study.

Previous work looking into uncertainty in the input for tsunami simulations often has focused on the earthquake source but mainly as a way of constraining the various proposed earthquake scenarios only after the event (see for example \cite{MacInnes:2013cr}).  In this work we concentrate on the uncertainty in the friction parameterization used in tsunami simulations.  Changes to the friction parameterization has been shown in the past to lead to significant changes to inundation levels \cite{Myers:2001el,Jakeman:2010hk} but these previous studies were limited in the scope of the changes allowed.  Recently, Sarri \emph{et al.} built a statistical surrogate of a landslide-generated tsunami computer model by using an emulator~\cite{Sarri2012}. The implemented emulator was the Gaussian process
that requires using a combination of prior knowledge about the simulator and
appropriate choices of functions and parameters. Their work was presented as a proof-of-concept case study
where they performed basic statistical and sensitivity analysis

In this work, we take advantage of the recent developments in Uncertainty Quantification 
(UQ) methods, which make it feasible to probe the sensitivity of complex and realistic tsunami models 
to uncertain parameters, and to assess the amount of information that can
be gained from specific measurements and observations. In particular, 
we implement an inverse modeling approach towards the objective stated above.  
Specifically, we will rely on Polynomial Chaos (PC) expansions to construct a 
faithful surrogate of the response of the large and complex geophysical 
model simulations.  The availability of the surrogate enables us to efficiently 
implement a Bayesian inference formalism to the inverse problem, namely 
because it dramatically reduces the cost of sampling the posterior distribution.

PC expansions have been developed and applied in the engineering community to quantify uncertainties
in numerical simulations; their principal use is in propagating input
uncertainties through large, complex and non-linear models to compute the
ensuing output uncertainties. Several researchers have been 
investigating the applicability of PC expansions for UQ in oceanic simulations
\citep{thacker2012,ashwanth2010,Alexanderian2012}. Recently,
PC combined with Bayesian inference was used to estimate drag coefficient
during Typhoon Fanapi 2010. The inference was done using
adjoint method and MCMC \cite{sraj:2013a,sraj:2013b}.

In this work, we follow a similar procedure as in \cite{sraj:2013a,sraj:2013b}
to estimate Manning N friction. We use the widely used Manning's-n law to provide a relationship between the friction coefficient and the roughness parameter $n$.  The parameter $n$ is also allowed to vary in space determined by the initial water depth to study effects of deep-water, near-shore, and on-shore sensitivities.  To carry out the uncertainty quantification we have employed polynomial chaos expansions.  Here the quantity of interest is taken to be the surface elevation $\eta$ and will compared to available NOAA Deep-ocean Assessment and Reporting of Tsunamis (DART) buoys for the \tohoku tsunami.  The forward model runs are performed using \geoclaw following the setup and using one of the source models explored in \cite{MacInnes:2013cr}.

>>>>

Tsunamis in the past decade have been responsible for some of the most 
deadly and costly natural disasters ever recorded. Coastal communities 
have faced this hazard by assessing the risk of these hazards, attempting 
to make informed decisions about the likelihood that such an event would 
occur with cost to life and property.  To do this, computational models of 
tsunami events are often employed to explore multiple different scenarios
to predict the water elevation caused by the tsunamis. The accurate prediction 
of water elevations requires accurate estimation of many model parameters that 
are either measured directly, defined empirically, or inferred from
sparsely collected data. Unfortunately, since tsunami events are relatively rare, 
there is an abundance of uncertainty in the input data for these computational 
models ranging from effects of the domain such as bathymetry and friction parameterizations, 
to the earthquake source.  In this article, we aim to estimate the Manning's n friction
coefficient that plays a key role in predicting the water elevation. 
Our approach is to quantify the uncertainty in the predicted water elevation 
and employ an inverse modeling approach to estimate the Manning's n coefficient 
using real observations of water elevation measured during the \tohoku tsunami as a case study.

Previous work looking into uncertainty in the input for tsunami simulations often has focused on the earthquake source but mainly as a way of constraining the various proposed earthquake scenarios only after the event (see for example \cite{MacInnes:2013cr}).  For instance, Sarri \emph{et al.} recently built a statistical surrogate of a landslide-generated tsunami computer model by using an emulator~\cite{Sarri2012}. The implemented emulator was the Gaussian process that requires using a combination of prior knowledge about the simulator and
appropriate choices of functions and parameters. Their work was presented as a proof-of-concept case study
where they performed basic statistical and sensitivity analysis. In this work we focus on the uncertainty in the friction parameterization used in tsunami simulations.  Changes to the friction parameterization has been shown in the past to lead to significant changes to inundation levels \cite{Myers:2001el,Jakeman:2010hk} but these previous studies were limited in the scope of the changes allowed.  In a recent work, Mayo et el. reformulated a statistical data assimilation method generally used in the estimation of model states to estimate Manning's n coefficient~\cite{Mayo2013}. They used a low-dimensional representations of Manning's n coefficients and recovered it by assimilating water elevation data.

In this work, we aim at making use of the recent developments in Uncertainty Quantification 
(UQ) methods that allows probing the sensitivity of realistic tsunami models 
to uncertain parameters, and inferring those parameters from specific observations. 
In particular, we implement an inverse modeling approach that relied on the Bayesian inference approach.  
We will also rely on Polynomial Chaos (PC) expansions to construct a 
faithful surrogate of the response of the large and complex geophysical 
model simulations.  The construction of the surrogate enables us to efficiently 
implement the Bayesian inference formalism as it dramatically reduces the cost of 
sampling the posterior distribution.

PC expansions have been developed and applied in the engineering community to quantify uncertainties
in numerical simulations; their principal use is in propagating input
uncertainties through large, complex and non-linear models to compute the
ensuing output uncertainties. Several researchers have been 
investigating the applicability of PC expansions for UQ in oceanic simulations
\citep{thacker2012,ashwanth2010,Alexanderian2012}. Recently,
PC combined with Bayesian inference was used to estimate drag coefficient
during Typhoon Fanapi 2010. The same problem was solved using
the adjoint method and MCMC \cite{sraj:2013a,sraj:2013b}.

In this work, we follow a similar procedure as in \cite{sraj:2013a,sraj:2013b}
to estimate Manning'n n coefficient of friction. 
%We use the widely used Manning's n law to provide a relationship between the friction coefficient and the roughness parameter $n$.  The parameter $n$ is also allowed to vary in space determined by the initial water depth to study effects of deep-water, near-shore, and on-shore sensitivities.  
To carry out the uncertainty quantification we have employed polynomial chaos expansions.  Here the quantity of interest is taken to be the surface elevation $\eta$ and will compared to available NOAA Deep-ocean Assessment and Reporting of Tsunamis (DART) buoys for the \tohoku tsunami.  The forward model runs are performed using \geoclaw following the setup and using one of the source models explored in \cite{MacInnes:2013cr}.

\alert{maybe an outline for this paper}

### 2013-12-23 paper/paper.tex
method section is completed for now
Minor modifications, mostly adding convenient packages
Fleshing out approach section
results section in progress and minor edits
Paper update
Minor changes to tsunami section
some edits to tsnunami section and adding authors
modified intro and added one section on manning

\usepackage{graphicx}
\usepackage{bm}
\usepackage{natbib}

% Add line numbers
\usepackage[mathlines]{lineno}

% Multipart figures
\usepackage{subcaption}

@@

\usepackage{ifpdf}

\usepackage{latexsym,amssymb,amsmath,color}
\newcommand{\revised}[1]{\textcolor{blue}{{#1}}}
\newcommand{\comment}[1]{\textcolor{red}{{#1}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\xxi}{\vec{\xi}}
\newcommand{\xxii}{\xxi^{(\scriptscriptstyle{-}\scriptstyle{i})}}
\newcommand{\var}[1]{{\mathrm{Var}}\left[ {#1} \right]}
\newcommand{\ip}[2]{\left( {#1}, {#2} \right)}
\newcommand{\U}{\mathcal{U}}
\newcommand{\normim}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega^{*})}}
\newcommand{\avemu}[1]{\mathrm{E}\left({#1}\right)}
\newcommand{\ave}[1]{\left\langle {#1} \right\rangle}
\newcommand{\prob}[1]{\mathrm{Prob}\left\{ {#1} \right\}}
\newcommand{\ind}[1]{\mathrm{\chi}_{\scriptscriptstyle {#1} }}
\newcommand{\NISP}{\mathcal{S}}
\newcommand{\ipmu}[2]{\left( {#1}, {#2} \right)_\mu}
\newcommand{\norm}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega)}}
\newcommand{\pard}[2]{\frac{\partial{#1}}{\partial{#2}}}
\def\xbold{{\vec{x}}}
\renewcommand{\vec}[1]{{\mathchoice
                     {\mbox{\boldmath$\displaystyle{#1}$}}
                     {\mbox{\boldmath$\textstyle{#1}$}}
                     {\mbox{\boldmath$\scriptstyle{#1}$}}
                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}

\newcommand{\geoclaw}{{\sc GeoClaw}\xspace}
\newcommand{\tohoku}{T\={o}hoku\xspace}

@@

\title{Polynomial Chaos for the Estimation of Manning's Friction}
\date{\today}

\author[kaust]{Ihab Sraj}
\ead{ihab.sraj@kaust.edu.sa}
\ead{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\author[ut]{Kyle T. Mandli}
% \ead{kyle@ices.utexas.edu}
% \ead[http://users.ices.utexas.edu/~kyle]{http://users.ices.utexas.edu/~kyle}
\author[kaust]{Ibrahim Hoteit}
\author[ut]{Clint N. Dawson}

\address[kaust]{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\address[ut]{Institute for Computational Engineering and Science, University of Texas at Austin, 201 E 24th ST. Stop C0200, Austin, TX 78712-1229, USA}

\date{\today}

@@

@@

\begin{keyword}
tsunami \sep Manning's N friction \sep sensitivity analysis \sep polynomial chaos
\end{keyword}

@@

\input{results}

\input{figures} 
\clearpage

>>>>

\usepackage{graphicx}
\usepackage{bm}
\usepackage{natbib}
\usepackage{xspace} % Incorporates correct optional space after commands
\usepackage{hyperref}

% Add line numbers
\usepackage[mathlines]{lineno} % Control line numbers

% Multipart figures
\usepackage{subcaption}  % "Newest" subfigure like package

@@

\usepackage{ifpdf} % For declaring rules for graphics below

\usepackage{latexsym,amssymb,amsmath,color}
\newcommand{\revised}[1]{\textcolor{blue}{{#1}}}
\newcommand{\comment}[1]{\textcolor{red}{{#1}}}
\newcommand{\alert}[1]{\textbf{\textcolor{red}{{#1}}}}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\xxi}{\vec{\xi}}
\newcommand{\xxii}{\xxi^{(\scriptscriptstyle{-}\scriptstyle{i})}}
\newcommand{\var}[1]{{\mathrm{Var}}\left[ {#1} \right]}
\newcommand{\ip}[2]{\left( {#1}, {#2} \right)}
\newcommand{\U}{\mathcal{U}}
\newcommand{\normim}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega^{*})}}
\newcommand{\avemu}[1]{\mathrm{E}\left({#1}\right)}
\newcommand{\ave}[1]{\left\langle {#1} \right\rangle}
\newcommand{\prob}[1]{\mathrm{Prob}\left\{ {#1} \right\}}
\newcommand{\ind}[1]{\mathrm{\chi}_{\scriptscriptstyle {#1} }}
\newcommand{\NISP}{\mathcal{S}}
\newcommand{\ipmu}[2]{\left( {#1}, {#2} \right)_\mu}
\newcommand{\norm}[1]{\left\| {#1} \right\|_{\scriptscriptstyle L^{2}(\Omega)}}
\newcommand{\pard}[2]{\frac{\partial{#1}}{\partial{#2}}}
\def\xbold{{\vec{x}}}
\renewcommand{\vec}[1]{{\mathchoice
                     {\mbox{\boldmath$\displaystyle{#1}$}}
                     {\mbox{\boldmath$\textstyle{#1}$}}
                     {\mbox{\boldmath$\scriptstyle{#1}$}}
                     {\mbox{\boldmath$\scriptscriptstyle{#1}$}}}}

\newcommand{\pd}{\pard}
\newcommand{\geoclaw}{{\sc GeoClaw}\xspace}
\newcommand{\tohoku}{T\={o}hoku\xspace}
\newcommand{\dx}{\ensuremath{\Delta x}}   % Delta x
\newcommand{\dy}{\ensuremath{\Delta y}}   % Delta y
\newcommand{\dt}{\ensuremath{\Delta t}}   % Delta t

@@

\title{Estimation of Manning's Friction Coefficient using Polynomial Chaos and Bayesian inference}
\date{\today}

\author[kaust]{Ihab Sraj}
\ead{ihab.sraj@kaust.edu.sa}
\ead{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\author[ut]{Kyle T. Mandli}
% \ead{kyle@ices.utexas.edu}
% \ead[http://users.ices.utexas.edu/~kyle]{http://users.ices.utexas.edu/~kyle}
\author[duke]{Omar M. Knio}
\author[ut]{Clint N. Dawson}
\author[kaust]{Ibrahim Hoteit}

\address[kaust]{Department of Physical Sciences and Engineering, King Abdullah University for Science and Technology, Thuwal, Saudi Arabia}
\address[ut]{Institute for Computational Engineering and Science, University of Texas at Austin, 201 E 24th ST. Stop C0200, Austin, TX 78712-1229, USA}
\address[duke]{Department of Mechanical Engineering and Materials Science, Duke University, 144
Hudson Hall, Durham, North Carolina 27708, USA}
\date{\today}

@@

Tsunamis have been lately catastrophic due to the warnings
issued late and to poorly predicted water elevations. Computational 
models of tsunami events are often employed to predict water elevations 
by exploring multiple different scenarios due to the abundance of uncertainty 
in the input data. Accurate estimation of water elevations requires 
accurate estimation of many model parameters. Our objective here is to 
present an efficient method to estimate the Manning's n friction coefficient
that plays an important role in the tsunami governing equations. 
In our work, we perform a sensitivity analysis of the Manning’s n coefficient 
and infer using water surface elevation data obtained during \tohoku tsunami. 
We characterize the Manning’s friction coefficient by three different parameters 
assumed constant in three regions: $n_1$ on-shore, $n_2$ near-shore and $n_3$ deep-water. 
The efficiency of our approach stems from the use of polynomial chaos expansions to build an inexpensive surrogate for the numerical tsunami \geoclaw model that can be used to perform the sensitivity analysis. The surrogate also reduces the computational burden of the Markov Chain Monte-Carlo sampling needed for the Bayesian inference. Our objective is to sharpen three initial estimates of the three uncertain parameters. Our results indicate that Manning’s n friction coefficients have a Maximum-A-Posteriori (MAP) values of $n_2=0.011$ and $n3=0.185$ while for $n_1$ no meaningful MAP value can be determined using the available data.

@@

\begin{keyword}
tsunami \sep Manning's n friction coefficient \sep sensitivity analysis \sep polynomial chaos \sep Bayesian inference
\end{keyword}

@@

\input{results}
\clearpage

\input{figures}

### 2013-12-23 paper/tsunami.tex
modified intro and added one section on manning

%!TEX root = paper.tex

@@

The observations used in this study are from the Deep-ocean Assessment and Reporting of Tsunamis (DART) buoy system developed and maintained by the National Oceanic and Atmospheric Administration (NOAA).  The purpose of the network is to provide early-warning detection and forecasting of tsunami propagation in the Pacific Ocean \alert{need citation for this?}.  The DART buoys closest to the earthquake source of the \tohoku tsunami were buoys 21401, 21413, 21418, and 21419 whose locations are shown in Figure~\ref{fig:setup}(left) and whose de-tided data for the event are shown in Figure~\ref{fig:setup}(right).
\subsection{Parametric Uncertainty}
In this work, we seek to quantify parametric uncertainty namely due 
to the uncertainty in the roughness parameter $n$. To this end,
we characterize the roughness $N$ by three different parameters
that are varied in three different regions but are assume
constant in each region. The three regions are as follows: 
on-shore who coefficnet is denoted by $N_1$, near-shore 
denoted by $N_2$, and deep-water denoted by $N_3$.
The three regions are shown in Figure~\ref{fig:ceofs}.
To carry out the uncertainty quantification we have employed polynomial chaos expansions 
as used in \cite{sraj:2013a,sraj:2013b}. The method and its application
to our problem is described in Section~\ref{sec:uqpce}.  
To represent the uncertainty in the Manning's $N$ coefficients a uniform distribution was assumed as the prior where $N$ was sampled in the interval $N \in [0.005-0.2]$.

>>>>

\section{Manning’s Friction Coefficient}
%We use the widely used .  The parameter $n$ is also allowed to vary in space determined by the initial water depth to study effects of deep-water, near-shore, and on-shore sensitivities.

Manning's $n$ law provides a relationship between the friction coefficient appearing in the shallow
water equations governing tsunamis and the roughness parameter $n$. 
This Manning's $n$ coefficient is variable that is highly spatial 
and dependent on the surface characteristics of the sea-floor. 
The direct measurement of $n$ is therefore impossible, instead, values
are often assigned from tables that are empirically estimated
for common types of land that are determined from satellite images.
As a result, the Manning’s n coefficients used in tsunami models
often contain large amounts of uncertainty.

The uncertainty in the Manning's $n$ coefficient leads to uncertainties
in the predicted quantities of interest such as the water elevations
and becomes a health risk during tsunami events. In addition, 
the prediction of these quantities of interest are often highly sensitive to 
the changes in these uncertain parameters. 
In this work, we seek to quantify parametric uncertainty namely due 
to the uncertainty in the roughness parameter $n$. To this end,
we characterize the roughness $n$ by three different parameters
that are varied in three different regions but are assumed
constant in each region. The three regions are as follows: 
on-shore whose coefficient is denoted by $n_1$, near-shore 
denoted by $n_2$, and deep-water denoted by $n_3$.
The three regions are shown in Figure~\ref{fig:ceofs}.
To carry out the uncertainty quantification we have employed polynomial chaos expansions 
as used in \cite{sraj:2013a,sraj:2013b}. The method and its application
to our problem is described in Section~\ref{sec:uqpce}.  
To represent the uncertainty in the Manning's $n$ coefficients a uniform distribution was assumed as the prior where $N$ was sampled in the interval $n \in [0.005-0.2]$.

@@

The observations used in this study are from the Deep-ocean Assessment and Reporting of Tsunamis (DART) buoy system developed and maintained by the National Oceanic and Atmospheric Administration (NOAA).  The purpose of the network is to provide early-warning detection and forecasting of tsunami propagation in the Pacific Ocean \alert{need citation for this?}.  The DART buoys closest to the earthquake source of the \tohoku tsunami were buoys 21401, 21413, 21418, and 21419 whose locations are shown in Figure~\ref{fig:setup}(left) and whose de-tided data for the event are shown in Figure~\ref{fig:setup}(right).
